Burning sage: Reversing the curse of dimensionality in the visualization of high-dimensional data

### 2021-01-25 paper/paper.tex
Updated tex file.

\textcolor{red}{The underlying idea of the (radial) transformation, to reverse the curse, is to correct for the distortion in geometry that occurs when projecting from a high-dimensional space onto low dimensions. This is done by mapping equal volume in the full space onto equal area in a two-dimensional projection via a geometrical transformation of the projected data. When used with data that is uniformly distributed in the high-dimensional space, the transformation will result in a unform distribution in the low-dimensional display. In general, for \textit{any} underlying data distribution, we can instead think about the effect in terms of resolution: adjusting for the geometric piling effect means that the resolution depends on the projected volume. The observed projected distribution is representative of the original high-dimensional distribution in this sense.}

@@

\textcolor{red}{The third common pre-processing would involve a dimension reduction step. Even with a relatively low number of dimensions, e.g. example with $p=10$, we see noticeable piling. However, we are often confronted with much higher dimensional datasets in which many of the variables are primarily adding noise. In this case it is advisable to first reduce the number of dimensions, for example using principal components analysis. Other dimension reduction techniques could be used. The sage tour would then be used to view the first few components, which would likely provide more insight about the data structures than is possible from only the first two. }

@@

Data were obtained from CT14HERA2 parton distribution function fits and
describe the sensitivity of fit parameters to experimental measurements
\citep{Wang:2018heo}. There are 28 parameters, and varying one at a time
to move \(\pm 1 \sigma\) away from the `best fit point' (maximum
likelihood estimate) provides our input variables, labelled X1-X56. Each
of the 2808 observations corresponds to a physical observable and
measures how the fit prediction changes along the 56 directions in
parameter space. Points are grouped based on the underlying process in
the experiment, which is mapped to color in the following. With the
analysis of the distribution along these variables X1-X56 we can
understand to what extent each experimental measurement provides new
information for the global fit. For example, orthogonality between
groups marks complementary constraints, and outlying points are
considered as important for future fits, see discussion in
\citet{Cook:2018mvr}.

>>>>

\textcolor{red}{The underlying idea of the (radial) transformation, to reverse the curse, is to correct for the distortion in geometry that occurs when projecting from a high-dimensional space onto low dimensions. This is done by mapping equal volume in the full space onto equal area in a two-dimensional projection via a geometrical transformation of the projected data. When used with data that is uniformly distributed in the high-dimensional space, the transformation will result in a uniform distribution in the low-dimensional display. In general, for \textit{any} underlying data distribution, we can instead think about the effect in terms of resolution: adjusting for the geometric piling effect means that the resolution depends on the projected volume. The observed projected distribution is representative of the original high-dimensional distribution in this sense.}

@@

\textcolor{red}{The third common pre-processing would involve a dimension reduction step. Even with a relatively low number of dimensions, e.g. with $p=10$, we see noticeable piling. However, we are often confronted with much higher dimensional datasets in which many of the variables are primarily adding noise. In this case it is advisable to first reduce the number of dimensions, for example using principal components analysis. Other dimension reduction techniques could be used. The sage tour would then be used to view the first few components, which would likely provide more insight about the data structures than is possible from only the first two. }

@@

Data were obtained from CT14HERA2 parton distribution function fits and
describe the sensitivity of fit parameters to experimental measurements
\citep{Wang:2018heo}. There are 28 parameters, and varying one at a time
to move \(\pm 1 \sigma\) away from the `best fit point' (maximum
likelihood estimate) provides our input variables, labeled X1-X56. Each
of the 2808 observations corresponds to a physical observable and
measures how the fit prediction changes along the 56 directions in
parameter space. Points are grouped based on the underlying process in
the experiment, which is mapped to color in the following. With the
analysis of the distribution along these variables X1-X56 we can
understand to what extent each experimental measurement provides new
information for the global fit. For example, orthogonality between
groups marks complementary constraints, and outlying points are
considered as important for future fits, see discussion in
\citet{Cook:2018mvr}.
